---
title: "betty"
author: "Betty Krebs"
format: html
---

```{r}
here::i_am("Womenempowerment.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
```

#Initialisation : 
Converting `mpi`from .xlsx to .csv
```{r}
library(readxl)
GMPI <- read_excel("~/dm-final-project/mpi.xlsx")
write.csv(GMPI, "C:/Users/bkreb/Documents/mpi.csv", row.names = FALSE)
GMPI_vroom <- vroom("~/dm-final-project/mpi.csv")
```
##Dataset Empowerment 

###Short description of the source

This dataset also comes from the World Health Organization, which issues the
survey-based women's empowerment index (SWPER). 
From this survey are drawn 9 indicators,  distributed across 3 domains : 
-the *social independence domain* : women are classified according to their empowerment ie women with high, medium or low empowerment in the social independence domain
-the *decision-making domain* : the dataset determines women with high empowerment in this domain, and with medium or low empowerment.
-the *violence domain* : women are again classified whether they benefit from high, medium or low empowerment in this domain. 

Those 9 indicators are disaggregated by age (with two subgroups 15-19yo and 20-49yo), economic status (determined within the country by calculating quintiles and deciles of the population), education (highest level of education reached by the women), place of residence and subnational region.

###Table which gives basic information about the data file

```{r}

```


###The most important variables 

##Description of the data joining and cleaning 

##Description of our research question 

##Graphical representation of our main target variable 