---
title: "Finalphaseproject"
author: "Betty Krebs & Marilys Guiban"
format: pdf
filters:
  - wordcount
editor: 
  markdown: 
    wrap: 72
---

##Initialisation

```{r}
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(openxlsx)
library(knitr)
```

## Research question presentation

Research question : Do the quality of institutions and the development
of institutional financing systems enable women to empower themselves in
the social and decision-making domains ?

The aim of our research question is to understand whether countries with
higher-governance and more developed financial inclusion infrastructures
provide better opportunities for women to participate in economic,
social and decision-making activities. By combining multiple datasets,
our study integrates information on financial behavior, governance
indicators and women's empowerment to explore this relationship in a
comprehensive manner.

Dependent variables : women empowerment in the social domain and in the
decision making domain

Explanatory variables : From the Findex dataset : borrowing from family
and friends (borrow_family_friends), using digital payments
(digital_payment), saving in a bank account (save_bank_account),
borrowing to start a business (borrow_start_business), and having a bank
account (has_bank_account)

From the Worldwide governance index : Voice and Accountability: Estimate
and Voice and Accountability: Percentile Rank, Rule of Law: Estimate and
Rule of Law: Percentile Rank, and Political Stability and Absence of
Violence/Terrorism: Estimate and Political Stability and Absence of
Violence/Terrorism: Percentile Rank

Our analysis relies primarily on three datasets. First the Global Findex
Database 2025, which privides demand-side data on financial inclusion,
including indicators such as account ownership, saving behaviour,
borrowing formal and informal sources and use of digital financial
services. These variables are critical to understanding women's access
to financial resources and their ability to manage risk and participate
in economic activity.

The WGI dataset offers a measure of institutional quality across
multiple dimensions such as political stability, rule of la, regulatory
quality and voice and accountability. These indicators allow us to
assess the institutional context in where financial systems operate.

The Empowerment dataset.

To provide a more nuanced analysis, we will do distinctions between
countries to see how cultural and legal frameworks often influence
women's access to financial resources and participation in economic
activity.

In term of methodology our project will combine descriptive statistics,
correlation analysis and regression models. The goal is to assess the
relationship between governance quality, financial inclusion and women's
empowerment. The clean datasets allow us to focus on the most relevant
variables.

In our analysis we will focus on the most recent data and give historical analysis with previous litterature made on this question. 

Pour répondre à cette question de recherche nous allons : 1° analyse
géographique comparative des pays par groupes de revenus et de
l'émancipation des femmes : par la réalisation de nuages de points
L'analyse se fera par : niveau de développement économique (middle, low,
high income countrie), par religion dominante (muslim vs non muslim en
ajoutant un nouveau dataset), 2° analyse de la corrélation entre finance
institutions et autonomie des femmes,

3° réaliser des régressions économétriques en isolant inclusion
financière et degré d'autonomie des femmes Les tests se composeront de
test de significativité des coefficients et de la qualité du modèle (à
réaliser sur plusieurs variables : 3 tests biens ciblés à choisir)

## Dataset description

### Empowerment

```{r}
Empowerment <- readRDS(here("datafinalproject", "Empowerment_clean.rds"))
```

This dataset comes from the World Health Organization, which issues the
survey-based women's empowerment index (SWPER). This index was first
tested in Africa in 2017 and was then extended to the rest of the world.
It is the first index in the gender domain to propose an
individual-level comparison between a large number of countries and over
time.

From this survey are drawn 9 indicators, distributed across 3 domains
: - *Social Independence Domain* - *Decision-Making Domain* - *Violence
Domain*

### Findex (cleaned dataset)

```{r}
Findex <- readRDS(here("datafinalproject", "Findexclean.rds"))
```

In the cleaned format of this dataset we have 8 variables : - Country,
codecountry and year : giving informations about the observations

-   "has_bank_account" : The percentage of respondents who report being
    employed in the private sector and receiving any money from their
    employer in the past year in the form of a salary or wages for doing
    work, and who received that money directly into a bank or similar
    financial institution account

-   "borrow_start_business" : The percentage of respondents who report
    borrowing any money to start or operate a business in the past year.

-   "save_bank_account" : The percentage of respondents who report
    saving or setting aside any money at a bank or similar financial
    institution or using a mobile money account to save in the past
    year.

-   "digital payment" : The percentage of respondents who report using
    mobile money, a debit or credit card, or a mobile phone to make a
    payment from an account--or report using the internet to pay bills
    or to buy something online or in a store--in the past year. This
    includes respondents who report paying bills, sending or receiving
    remittances, receiving payments for agricultural products, receiving
    government transfers, receiving wages, or receiving a public sector
    pension directly from or into a bank or similar financial
    institution account or through a mobile money account in the past
    year.

-   "borrow family_friends" : The percentage of respondents who report
    borrowing any money from family, relatives, or friends in the past
    year

### WGI (cleaned format)

```{r}
WGI <- readRDS(here("datafinalproject", "WGI_cleaned.rds"))
```

For our research question we are using : - political stability and the
absence of terrorism as an explicative variable of women empowerment It
measures perceptions of the likelihood of political instability and/or
politically-motivated violence, including terrorism. Estimate gives the
country's score on the aggregate indicator, in units of a standard
normal distribution, i.e. ranging from approximately -2.5 to 2.5.

-   Voice and accountability : captures perceptions of the extent to
    which a country's citizens are able to participate in selecting
    their government, as well as freedom of expression, freedom of
    association, and a free media. Estimate gives the country's score on
    the aggregate indicator, in units of a standard normal distribution,
    i.e. ranging from approximately -2.5 to 2.5.

-   rule of law : Regulatory Quality captures perceptions of the ability
    of the government to formulate and implement sound policies and
    regulations that permit and promote private sector development.

For each indicators we keep the estimate and the percentile value: -
Estimate gives the country's score on the aggregate indicator, in units
of a standard normal distribution. - Percentile rank indicates the
country's rank among all countries covered by the aggregate indicator,
with 0 corresponding to lowest rank, and 100 to highest rank. Percentile
ranks have been adjusted to correct for changes over time in the
composition of the countries covered by the WGI.

### Joined file

The final dataset integrates all variables from the aforementioned
sources. By using the country code and year as primary keys, we
successfully merged the data into a single, comprehensive table for our
analysis.

```{r}
final_dataset <- readRDS(here("datafinalproject", "Final_Dataset.rds"))
```

To improve analytical precision, the final_recent dataset was restricted to 2023 empowerment values. This minimizes the time gap with the 2024 WGI and Findex data, ensuring that the relationship between governance and empowerment is based on nearly contemporaneous observations.
```{r}
extract_value <- function(x) {
  if (is.list(x) && length(x) > 0) {
    return(as.numeric(x[[1]]))
  } else if (is.numeric(x)) {
    return(x)
  } else {
    return(NA)
  }
}

final_cleaned_all <- final_dataset %>%
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) %>%
  mutate(
    # Application de ta fonction aux 3 colonnes WGI
    political_stability = sapply(`Political Stability and Absence of Violence/Terrorism: Percentile Rank`, extract_value),
    voice_accountability = sapply(`Voice and Accountability: Percentile Rank`, extract_value),
    rule_of_law = sapply(`Rule of Law: Percentile Rank`, extract_value),
    
    # Nettoyage du nom des domaines
    domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")
  )

final_recent <- final_cleaned_all %>%
  group_by(country, domaine) %>%
  filter(year == max(year)) %>%
  ungroup()
```

(la première partie à inclure dans la partie 2 pour plus de propreté)

## Graphical representations

In this part we will do graphical analysis to understand what is
significant in women's empowerment. We decided not to analyse high
income countries since our research question explores development issues
and also due to a lack of data.

### Analysis of empowerment by income group

```{r}
final_recent %>%
  # On enlève les lignes où l'un des deux est vide
  filter(!is.na(estimate), !is.na(wbincome2024), wbincome2024 != "High income") %>% 
  group_by(wbincome2024, country) %>% 
  summarise(score_empowerment = mean(estimate, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = wbincome2024, y = score_empowerment, fill = wbincome2024)) +
  geom_boxplot() + 
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal()
```
The distribution of empowerment scores reveals a significant shift across income groups. While Low-Income countries struggle with high inequality and low averages, Lower-Middle-Income nations show a consolidated progress with more homogeneous and higher scores. In contrast, Upper-Middle-Income countries exhibit a dual reality: despite a higher ceiling of empowerment, the low positioning of the mean within a tall box suggests persistent internal disparities, where a large segment of the female population has yet to catch up with the top achievers.

Let's analyse this deeper by focusing on social domain and decision
making domain by countries :

```{r}
#| warning: false
#| message: false

final_recent |>
  filter(indicator_name %in% c(
    "Women with high empowerment in the attitude to violence domain (%)",
    "Women with high empowerment in the decision-making domain (%)",
    "Women with high empowerment in the social independence domain (%)"
  ), wbincome2024 != "High income") |>
  group_by(country, wbincome2024, indicator_name) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  ggplot(aes(x = wbincome2024, y = score_moyen, fill = wbincome2024)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.7) +
  facet_wrap(~domaine, scales = "free_y") +
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

When disaggregating the data by income group, it is evident that
Upper-Middle-Income countries consistently exhibit a higher proportion
of empowered women across all three domains—Social, Decision-Making, and
Attitude toward Violence—compared to both Lower-Middle and Low-Income
countries.

### Analysis of empowerment by the quality of institutions

In order to do this we will analyse graphically whether there seem to be
a correlation between empowerment and the quality of institutions.

*Political stability and women empowerment*

```{r}
# Diagnostic de la colonne
class(final_dataset$`Political Stability and Absence of Violence/Terrorism: Percentile Rank`)
head(final_dataset$`Political Stability and Absence of Violence/Terrorism: Percentile Rank`, 3)
```


*Political stability and women empowerment*

```{r}
final_recent %>%
  group_by(country, wbincome2024, domaine, political_stability) %>%
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(political_stability), !is.na(score_moyen)) %>%
  ggplot(aes(x = political_stability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Does Political Stability Foster Women's Empowerment?",
    subtitle = "Comparison across economic levels (Low to Upper-Middle Income)",
    x = "Political Stability Index (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```
The divergent trends across income groups suggest that the relationship
between political stability and empowerment is not linear but
context-dependent. While stability is a fundamental prerequisite for
empowerment in Low-Income settings, its impact diminishes or even
reverses in Lower-Middle-Income contexts. In contrast, for
Upper-Middle-Income nations, the positive correlation in decision-making
indicates that stability provides the necessary environment for highly
educated women to consolidate their role in governance and household
management.

=\> AI advice : look at the authoritarian paradox

*Voice and accountability and Women empowerment*

```{r}
final_recent %>%
  group_by(country, wbincome2024, domaine, voice_accountability) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(voice_accountability), !is.na(score_moyen)) |>
  ggplot(aes(x = voice_accountability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Voice and Accountability vs Women's Empowerment",
    x = "Voice & Accountability (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal()
```
Voice and Accountability shows a positive correlation with reduced violence across most groups, yet reveals a striking divergence in autonomy: while Upper-Middle-Income countries see gains across all domains, Low and Lower-Middle-Income nations experience a 'participation paradox' where increased political voice coincides with a decline in domestic decision-making and social independence.

*Rule of law and women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, rule_of_law) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(rule_of_law), !is.na(score_moyen)) |>
  ggplot(aes(x = rule_of_law, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Rule of Law vs. Women's Empowerment",
    subtitle = "Analysis of legal frameworks across income groups",
    x = "Rule of Law (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```
Rule of Law acts as a universal driver for social independence and attitudes against violence across all income groups; however, it fails to penetrate the domestic sphere in lower-income economies, where decision-making power continues to decline despite strengthening legal frameworks.

### Analysis of empowerment by development of financing systems
```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, digital_payment) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(digital_payment), !is.na(score_moyen)) |>
  
  ggplot(aes(x = digital_payment, y = score_moyen, color = wbincome2024)) +
  # 1. Les points sont dimensionnés par le nombre de pays à cet endroit précis
  geom_count(alpha = 0.3, size = 1) + 
  # 2. Une ligne de tendance avec ruban d'incertitude (se = TRUE) pour voir la fiabilité
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  # 4. Couleurs mieux contrastées
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the development of digital payment on Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population with Digital Payments",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```
The data on digital financial inclusion reveals a complex landscape: while it serves as a catalyst for changing attitudes toward violence in Lower-Middle-Income countries, it shows a global negative correlation with decision-making and social independence, suggesting that digital tools alone—potentially due to data scarcity or the 'digital divide'—have yet to translate into broader structural autonomy for women in developing economies.


```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, has_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(has_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = has_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population which has a bank accont and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who has a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```
Analysis of formal bank account ownership reveals a transitional stage: while it fosters social independence and decision-making power in Lower-Middle-Income countries, it shows a slight negative correlation with attitudes toward violence globally and a decline in decision-making for Upper-Middle-Income groups, potentially reflecting a shift where basic banking is no longer a sufficient lever for empowerment in more advanced economies.

```{r}
final_cleaned_all |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, borrow_start_business) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(borrow_start_business), !is.na(score_moyen)) |>
  
  ggplot(aes(x = borrow_start_business, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population which borrow to start its business and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who borrow to start a business",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```
The data analysis reveals that while borrowing to bank to start a business correlate with increased social independence in Lower-Middle-Income countries, they show a global decline in decision-making power and attitudes toward violence, suggesting that digital financial tools may increase women's mobility without yet shifting deep-seated domestic power dynamics or social norms.


```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, save_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(save_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = save_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population who save on a bank account and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who saves of a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```
The analysis of formal savings shows that while it correlates with increased attitudes against violence in Lower-Middle-Income countries, it appears decoupled from broader empowerment; in fact, formal savings coincide with a decline in decision-making and social independence across most groups, suggesting that saving may be a protective financial strategy rather than a tool for immediate domestic or social autonomy.



