---
title: "Determinants of Women's Empowerment: A Cross-Country Analysis"
author: "Betty Krebs & Marilys Guiban"
format: 
  pdf:
    echo: false   
    message: false
    warning: false
    toc: true
    number-sections: true
    colorlinks: true
    geometry:
      - margin=1in
editor: 
  markdown: 
    wrap: 72
---

*This report was generated using Quarto and RStudio. All statistical
analyses were performed in R, and the final document was compiled into
PDF format using the TinyTeX distribution to ensure high-quality
typesetting of econometric tables and LaTeX formulas.*

```{r}
# install.packages("WDI")  
# install.packages("sandwich") 
# install.packages("lmtest")
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(knitr)
library(stringr)
library(WDI)
library(lmtest)
library(sandwich)
library(car)
```

## Research question presentation

Women empowerment is addressed by the United Nations as a key step towards development as it holds the fifth position among the Sustainable Development Goals (SDGs), just behind the objectives of no poverty, zero hunger, good health and quality education. 

Yet, as of January 2025, women held only 27.2% of national parliament seats. This represents a tiny 0.3 point increase compared to 2024. 

Recent literature shows that the progress is even slower in the financial domain, thus hindering women's empowerment. Scholars stress the need for more financial independence for women. For instance, Sujan (2016) makes a list of financial tasks essential in a household that could be endorsed by women, Bhumija Chauhan (2024) shows there is a robust evidence of a correlation between financial independence and economic, social and political empowerment. Pal, Gupta and Joshi(2022) show that earning status, social security schemes, and financial decision-making at household level are significant to promote women empowerment through financial inclusion. 

There is a significant gap regarding how institutional quality and financial infrastructure interact to produce empowerment. Specifically, we argue that individual access to bank accont may be insufficient if the rule of law is weak or if political instability restricts her mobility. 

Our research aims at studying the interplay between institutional quality, legal frameworks, and financial infrastructure.

Research question : Do the quality of institutions and the development
of institutional financing systems enable women to empower themselves in
the social, decision-making and ability to respond to violence domains ?


**Variable selection** 
The objective of our study is to determine whether countries with higher-governance and mature financial inclusion infrastructures
provide better opportunities for women to participate in economic, social and decision-making activities. 

*Dependent variables* : we focus on three dimensions of the WHO's survey-based Women's Empowerment (SWPER) index : the social independence and decision-making domain and the attitude of women toward violence. 

*Explanatory variables* 
Financial inclusion (*Findex 2025*) : borrowing from family
and friends (borrow_family_friends), using digital payments
(digital_payment), saving in a bank account (save_bank_account),
borrowing to start a business (borrow_start_business), and having a bank
account (has_bank_account)

From the Worldwide governance index (*WGI*) : Voice and Accountability, Rule of Law, Political Stability and Absence of violence or terrorism

Our analysis relies primarily on three datasets :

- Findex dataset (2025) which provides demand-side data on financial inclusion,
including indicators such as account ownership, saving behaviour. These variables are critical to understanding women's access
to financial resources and their ability to manage risk and participate
in economic activity.

- WGI dataset : it offers a measure of institutional quality across
multiple dimensions such as political stability, rule of law, regulatory
quality and voice and accountability. These indicators allow us to
assess the institutional context in where financial systems operate.

- The Empowerment dataset proposes an estimation of the empowerment of women in the social independence domain, the decision-making model and the attitude to violence.

**Methodology** 

1. **Cross-country analysis** : we examine in a first part scores across different income groups (GDP) and cultural contexts using muldimensional visualizations 

2.**Correlation analysis** : we identify correlations between financial infrastructure development and empowerment outcomes 

3. **Economic model and regression** : we estimate linear regression models with robustness tests to isolate the effects of institutions and finance, while using national wealth as a control variable (log(GDP)) to mitigate omitted variable bias. This approach allows us to tst the robustness of these determinants and identify potential paradoxes. 

## Dataset description

### Empowerment

```{r}
Empowerment <- readRDS(here("datafinalproject", "Empowerment_clean.rds"))
```

This dataset comes from the World Health Organization, which issues the
survey-based women's empowerment index (SWPER). This index was first
tested in Africa in 2017 and was then extended to the rest of the world.
It is the first index in the gender domain to propose an
individual-level comparison between a large number of countries and over
time.

From this survey are drawn 9 indicators, distributed across 3 domains:
    - *Social Independence Domain* 
    - *Decision-Making Domain*
    - *Violence Domain*

### Findex 

```{r}
Findex <- readRDS(here("datafinalproject", "Findexclean.rds"))
```

In the cleaned format of this dataset we have 8 variables : 

-   Country, codecountry and year : giving informations about the observations

-   "has_bank_account" : The percentage of respondents who report being
    employed in the private sector and receiving any money from their
    employer in the past year in the form of a salary or wages for doing
    work, and who received that money directly into a bank or similar
    financial institution account

-   "borrow_start_business" : The percentage of respondents who report
    borrowing any money to start or operate a business in the past year.

-   "save_bank_account" : The percentage of respondents who report
    saving or setting aside any money at a bank or similar financial
    institution or using a mobile money account to save in the past
    year.

-   "digital payment" : The percentage of respondents who report using
    mobile money, a debit or credit card, or a mobile phone to make a
    payment from an account--or report using the internet to pay bills
    or to buy something online or in a store--in the past year. This
    includes respondents who report paying bills, sending or receiving
    remittances, receiving payments for agricultural products, receiving
    government transfers, receiving wages, or receiving a public sector
    pension directly from or into a bank or similar financial
    institution account or through a mobile money account in the past
    year.

-   "borrow family_friends" : The percentage of respondents who report
    borrowing any money from family, relatives, or friends in the past
    year

### WGI (cleaned format)

```{r}
WGI <- readRDS(here("datafinalproject", "WGI_cleaned.rds"))
```

The main variables of this dataset are : 

-   political stability and the absence of terrorism as an explicative variable of women empowerment.      It measures perceptions of the likelihood of political instability and/or politically-motivated        violence, including terrorism. Estimate gives the country's score on the aggregate indicator, in       units of a standard normal distribution, i.e. ranging from approximately -2.5 to 2.5.

-   Voice and accountability : captures perceptions of the extent to
    which a country's citizens are able to participate in selecting
    their government, as well as freedom of expression, freedom of
    association, and a free media. Estimate gives the country's score on
    the aggregate indicator, in units of a standard normal distribution,
    i.e. ranging from approximately -2.5 to 2.5.

-   rule of law : Regulatory Quality captures perceptions of the ability
    of the government to formulate and implement sound policies and
    regulations that permit and promote private sector development.


For each indicators we keep the estimate and the percentile value: 

- Estimate gives the country's score on the aggregate indicator, in units
of a standard normal distribution. 

- Percentile rank indicates the country's rank among all countries covered by the aggregate indicator,
with 0 corresponding to lowest rank, and 100 to highest rank. Percentile
ranks have been adjusted to correct for changes over time in the
composition of the countries covered by the WGI.

### Joined file

The final dataset integrates all variables from the beforementioned
sources. By using the country code and year as primary keys, we
successfully merged the data into a single, comprehensive table for our
analysis.

```{r}
final_cleaned_all <- readRDS(here("datafinalproject", "final_cleaned_all.rds"))
```

To improve analytical precision, the final_recent dataset was restricted
to 2023 empowerment values. This minimizes the time gap with the 2024
WGI and Findex data, ensuring that the relationship between governanc, access to financial services
and empowerment is based on nearly contemporaneous observations.

```{r}
final_recent <- readRDS(here("datafinalproject", "final_recent.rds"))
```

```{r}
final_with_gdp <- readRDS(here("datafinalproject", "final_with_gdp.rds"))
```

```{r}
dataset_names <- c(
  "WHO Empowerment (SWPER)", 
  "Global Findex 2025", 
  "Worldwide Governance Indicators", 
  "Final Merged Dataset"
)

observations <- c(
  nrow(Empowerment), 
  nrow(Findex), 
  nrow(WGI), 
  nrow(final_recent)
)

variables <- c(
  ncol(Empowerment), 
  ncol(Findex), 
  ncol(WGI), 
  ncol(final_recent)
)


summary_df <- data.frame(
  "Dataset" = dataset_names,
  "Observations" = observations,
  "Variables" = variables
)


kable(summary_df, 
      align = "lcc", # Dataset à gauche (l), chiffres centrés (c)
      caption = "Table: Data structure across source and final datasets")
```



## Graphical representations

In this part we will do graphical analysis to understand what is
significant in women's empowerment. We decided not to analyse high
income countries since our research question explores development issues
and also due to a lack of data.

### Temporal analysis : how women's empowerment has evolved these past years ?

```{r}
final_cleaned_all |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  filter(!is.na(wbincome2024), wbincome2024 != "High income") |>
  group_by(year, wbincome2024, domaine) |> 
  summarise(mean_emp = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(mean_emp)) |> 
  ggplot(aes(x = year, y = mean_emp, color = wbincome2024, group = wbincome2024)) +
  geom_line(linewidth = 0.8, alpha = 0.6) + # 'linewidth' remplace 'size' pour les lignes
  geom_point(size = 1.5) + 
  facet_wrap(~domaine) +
  labs(
    title = "Historical Evolution of Women's Empowerment (2010-2023)",
    subtitle = "Trends by World Bank Income Group",
    x = "Year", 
    y = "Average Score (%)",
    color = "Income Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

The historical data (1999-2023) reveals an overall upward trend in
women's empowerment, particularly in the Attitude to Violence and
Decision-Making domains across all income groups. Notably, the Social Independence domain shows a concerning
stagnation, or even a slight decline in some Upper-Middle and Low-Income
contexts, indicating that physical mobility and financial autonomy
remain the most difficult barriers to break, despite improvements in
other areas.

Our temporal analysis reflect a shorter window of a longer historical process which was documented by Claudia Goldin (2006) in her study of the 20th century. In her U-shaped hypothesis, Claudia Goldin's explains the non-linear relationship between econimic development and women's labor participation to labor. In agrarian economies female labor is high because production takes place within the household and the family where women's work is seen as "socially-acceptable". As economies industrialize, their participation declines because they are spatially separated from the home, and industrial jobs are considered as too demanding for a married woman. In a thrid time, at higher level of develoment female labor participation increases as the economy shift toward service-sector. For Goldin it is in this part that fertility declines as they are empowering themselves by having their own source of income. Thus as it is observed in our graphics, female empowermetn does not automatically lead to continuous improvments in gender equality and the empowerment of women. 

In order to explain why women's empowerment in attitude toward violence seem to increase more significantly we could rely on Jaychandran (2015) study which shows that in order to allow women empowerment, it is harder to change cultural practices that are the main factors of women unempowerment. 

### Analysis of empowerment by income group

```{r}
final_recent |>
  filter(!is.na(estimate), !is.na(wbincome2024), wbincome2024 != "High income") |> 
  group_by(wbincome2024, country) |> 
  summarise(score_empowerment = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = wbincome2024, y = score_empowerment, fill = wbincome2024)) +
  geom_boxplot() + 
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal()
```

The distribution of empowerment scores reveals a significant shift
across income groups. While Low-Income countries struggle with high
inequality and low averages, Lower-Middle-Income nations show a
consolidated progress with more homogeneous and higher scores. In
contrast, Upper-Middle-Income countries exhibit a dual reality: despite
a higher ceiling of empowerment, the low positioning of the mean within
a tall box suggests persistent internal disparities, where a large
segment of the female population has yet to catch up with the top
achievers.

```{r}
final_recent |>
  filter(indicator_name %in% c(
    "Women with high empowerment in the attitude to violence domain (%)",
    "Women with high empowerment in the decision-making domain (%)",
    "Women with high empowerment in the social independence domain (%)"
  ), wbincome2024 != "High income") |>
  # We add a security filter pour empty scores
  filter(!is.na(estimate)) |> 
  group_by(country, wbincome2024, indicator_name) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  ggplot(aes(x = wbincome2024, y = score_moyen, fill = wbincome2024)) +
  # na.rm = TRUE suppresses the warning on the deleted lines
  geom_boxplot(outlier.shape = NA, alpha = 0.7, na.rm = TRUE) +
  facet_wrap(~domaine, scales = "free_y") +
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

A cross-sectional analysis by income group reveals heterogenous patterns
across the three empowerment domains. In the Attitude toward Violence
domain, a clear positive correlation with national wealth is observed as
the median score rises from 52% in Low-Income countries to 75% in
Upper-Middle-Income contexts. However, the widening of the interquartile
range in wealthier groups suggests that economic growth alone does not
harmonize social norms.

In contrast, the Decision-Making domain shows extreme volatility within
the Low-Income group, with scores ranging from 23% to 63%. This
'long-box' indicates that domestic authority can reach high
levels even in resource-constrained environments, likely driven by
specific institutional or cultural factors. Finally, Social Independence
appears to be the most rigid dimension, characterized by lower overall
scores and narrower distributions. The transition from a 23% median in
Low-Income countries to only 60% in Upper-Middle countries confirms that
physical and financial autonomy faces a 'structural ceiling' that is
significantly more resistant to economic development than other
dimensions of empowerment.

### Analysis of empowerment by the quality of institutions

In order to do this we will analyse graphically whether there seems to be
a correlation between empowerment and the quality of institutions.

*Political stability and women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, political_stability) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(political_stability), !is.na(score_moyen)) |>
  ggplot(aes(x = political_stability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Does Political Stability Foster Women's Empowerment?",
    subtitle = "Comparison across economic levels (Low to Upper-Middle Income)",
    x = "Political Stability Index (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```

The divergent trends across income groups suggest that the relationship
between political stability and empowerment is not linear but
context-dependent. While stability is a fundamental prerequisite for
empowerment in Low-Income settings, its impact diminishes or even
reverses in Lower-Middle-Income contexts. In contrast, for
Upper-Middle-Income nations, the positive correlation in decision-making
indicates that stability provides the necessary environment for highly
educated women to consolidate their role in governance and household
management.

*Voice and accountability and Women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, voice_accountability) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(voice_accountability), !is.na(score_moyen)) |>
  ggplot(aes(x = voice_accountability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Correlation between Voice and Accountability and Women's Empowerment",
    x = "Voice & Accountability (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal()
```

Voice and Accountability shows a positive correlation with reduced
violence across most groups, yet reveals a striking divergence in
autonomy: while Upper-Middle-Income countries see gains across all
domains, Low and Lower-Middle-Income nations experience a 'participation
paradox' where increased political voice coincides with a decline in
domestic decision-making and social independence.

*Rule of law and women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, rule_of_law) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(rule_of_law), !is.na(score_moyen)) |>
  ggplot(aes(x = rule_of_law, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Correlation between rule of law and Women's Empowerment",
    subtitle = "Analysis of legal frameworks across income groups",
    x = "Rule of Law (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```

Rule of Law acts as a universal driver for social independence and
attitudes against violence across all income groups; however, it fails
to penetrate the domestic sphere in lower-income economies, where
decision-making power continues to decline despite strengthening legal
frameworks.

### Analysis of empowerment by development of financing systems

*Digital payments*

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, digital_payment) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(digital_payment), !is.na(score_moyen)) |>
  
  ggplot(aes(x = digital_payment, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation development of digital payment vs Women's Empowerment",
    x = "% of population with Digital Payments",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The data on digital financial inclusion reveals a complex landscape:
while it serves as a catalyst for changing attitudes toward violence in
Lower-Middle-Income countries, it shows a global negative correlation
with decision-making and social independence, suggesting that digital
tools alone—potentially due to data scarcity or the 'digital
divide'—have yet to translate into broader structural autonomy for women
in developing economies.

*Is having a bank account correlated with women empowerment ?*

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, has_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(has_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = has_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation has a bank account vs Women's Empowerment",
    x = "% of population who has a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

Analysis of formal bank account ownership reveals a transitional stage:
while it fosters social independence and decision-making power in
Lower-Middle-Income countries, it shows a slight negative correlation
with attitudes toward violence globally and a decline in decision-making
for Upper-Middle-Income groups, potentially reflecting a shift where
basic banking is no longer a sufficient lever for empowerment in more
advanced economies.

*Ability to borrow to bank to start a business*

```{r}
final_cleaned_all |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, borrow_start_business) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(borrow_start_business), !is.na(score_moyen)) |>
  
  ggplot(aes(x = borrow_start_business, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation borrow to start its business vs Women's Empowerment",
    x = "% of population who borrow to start a business",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The data analysis reveals that while borrowing to bank to start a
business correlate with increased social independence in
Lower-Middle-Income countries, they show a global decline in
decision-making power and attitudes toward violence, suggesting that
digital financial tools may increase women's mobility without yet
shifting deep-seated domestic power dynamics or social norms.

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, save_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(save_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = save_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation save a bank account vs. Women's Empowerment",
    x = "% of population who saves of a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The analysis of formal savings shows that while it correlates with
increased attitudes against violence in Lower-Middle-Income countries,
it appears decoupled from broader empowerment; in fact, formal savings
coincide with a decline in decision-making and social independence
across most groups, suggesting that saving may be a protective financial
strategy rather than a tool for immediate domestic or social autonomy.

Limits of our analysis : we can observe that there is a lack of data on
low income country it could be due to


### Geographical analysis 


```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    )
  ) |>
  mutate(
    domaine = str_remove_all(
      indicator_name,
      "Women with high empowerment in the | domain \\(%\\)"
    )
  ) |>
  group_by(whoreg6, domaine) |>
  summarise(
    mean_empowerment = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = domaine, y = whoreg6, fill = mean_empowerment)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "C",
    name = "Mean empowerment (%)"
  ) +
  labs(
    title = "Average women's empowerment by region and domain",
    x = "Empowerment domain",
    y = "World Health Organisation region"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 20, hjust = 1),
    plot.title = element_text(face = "bold")
  )

```

```{r}
library(tidyr)

df_dumbbell <- final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    )
  ) |>
  mutate(
    domaine = str_remove_all(
      indicator_name,
      "Women with high empowerment in the | domain \\(%\\)"
    )
  ) |>
  group_by(whoreg6, domaine) |>
  summarise(
    mean_empowerment = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = domaine, values_from = mean_empowerment)

ggplot(df_dumbbell) +
  geom_segment(
    aes(
      y = whoreg6,
      yend = whoreg6,
      x = `attitude to violence`,
      xend = `decision-making`
    ),
    color = "grey70",
    linewidth = 1
  ) +
  geom_point(
    aes(x = `attitude to violence`, y = whoreg6),
    color = "#1b9e77",
    size = 3
  ) +
  geom_point(
    aes(x = `decision-making`, y = whoreg6),
    color = "#d95f02",
    size = 3
  ) +
  labs(
    title = "Gap between empowerment domains by WHO region",
    x = "Mean empowerment (%)",
    y = "WHO region"
  ) +
  theme_minimal()

```


## Econometric regression with scatter plots

In this part we are using a dataset including the GDP of each country and we created a region.

To compute our different tests, we will do 4 hypothesis : 

1° **the ability of women to empower themselves in the social sphere** depends on
political freedom and their ability to access formal financial services
To test this we will use : 

- as a dependent variable : women empowerment in the social domain 

- as an explicative variable : voice_accountability, has_bank_account, digital_payments and
log_gdp

2° **The ability of women to empower themselves in violent situations** depends on political stability, the power of law within a country and the ability to save money on a bank account  

To test this we will use 

- as a dependent variable : women attitude toward violence 

- as an explicative variable : log_gdp, political_stability, rule_of_law, save_bank_account

3° **The ability of women to empower themselves in the decision-making domain** 
To test this we will use : 

- as a dependent variable : women empowerment in the decision-making domain 

- as an explicative variable : log_gdp, voice_accountability, rule_of_law, borrow_start_business


**Test of women empowerment in social domain**

```{r}
fit <- lm(estimate ~ log_gdp + voice_accountability + digital_payment + has_bank_account, 
          data = final_with_gdp |> filter(domaine == "social independence"))

results_robust <- coeftest(fit, vcov = vcovHC(fit, type = "HC1"))

print(results_robust)
```

The empirical results reveal a complex relationship between economic
development, technology, and social empowerment:

-   **The Dominance of GDP in Development**: The coefficient for Log GDP
    per capita is positive and highly significant ( β =19.05,p<0.001).
    This confirms that national wealth remains the primary engine for
    social independence, likely by increasing access to education,
    infrastructure, and health services for women.

-   **The effect of freedom and democracy** : the access to freedom and especially free speech surprisingly does not allow for women empowerment in the social domain. On the contrary, it has a  significant negative effect on women empowerment (β=−0,124,p<0.01)

-   **The Digital Paradox**: The Digital Payment variable shows a
    significant negative coefficient (β=−22,736,p<0.05). This
    counter-intuitive result suggests that in the absence of targeted
    social policies, the mere expansion of digital financial tools does
    not automatically translate into greater social autonomy. It may
    reflect a "participation paradox" where digital tools are prevalent
    in societies where physical mobility remains restricted, or where
    digital surveillance offsets the benefits of financial inclusion.
    
Perhaps we could also explain that by saying that digital finance may increase women's economic activities without reducing domestic responsibilities, this may intensify time constraints and limiting their participation in social life. 

Another plausible explanation for these strange results may be that endogeneity or omitted variable bias related to unobserved social practices, which are not revealed within this dataset. 

**Attitude toward violence**

```{r}
fit_violence <- lm(estimate ~ log_gdp + rule_of_law + political_stability + save_bank_account, 
                   data = final_with_gdp %>% filter(domaine == "attitude to violence"))

results_violence <- coeftest(fit_violence, vcov = vcovHC(fit_violence, type = "HC1"))
print(results_violence)
```

This model examines the factors influencing women's rejection of domestic violence. The results, estimated with robust standard errors (HC1), highlight three primary dynamics : first, national wealth again (Log GDP) is a powerful and highly significant predictor (β=5.35,p<0.001), confirming that economic development is a fundamental driver for shifting social norms toward the rejection of violence. Second, formal financial inclusion through savings (save_bank_account) shows the strongest positive impact (β=20.08,p<0.001), suggesting that an independent "security buffer" significantly strengthens women's bargaining power and agency. Finally, the negative correlation with Rule of Law (β=−0.66,p<0.001) suggests a potential reporting bias: as legal frameworks strengthen and awareness grows, women may become more critical of violence and less likely to normalize it, leading to more conservative empowerment scores despite an improving institutional environment.


```{r}
fit_decision <- lm(estimate ~ log_gdp + voice_accountability + borrow_start_business, 
                   data = final_with_gdp %>% filter(domaine == "decision-making"))

results_decision <- coeftest(fit_decision, vcov = vcovHC(fit_decision, type = "HC1"))
print(results_decision)
```

This model reveals three different things : 
- *Positive Institutional Synergy* : Unlike the previous models, Voice and Accountability shows a positive and highly significant correlation with domestic decision-making (β=0.38,p<0.001). This suggests that political freedom and the ability to participate in the public sphere may act as a catalyst for women to claim more authority within the private domestic sphere.

- *The Entrepreneurial Credit Gap* : Interestingly, borrowing to start a business (borrow_start_business) does not have a statistically significant impact on decision-making power (β=1.88,p>0.1). As argued by Bernhardt et al. (2019), financial resources targeted at women are often redirected toward their husbands' activities. This implies that while a loan is officially registered to a woman, the economic returns and the subsequent power of decision are captured by the male spouse. Consequently, access to credit alone, without addressing the intra-household power dynamics, may fail to enhance women's domestic agency.

### Testing model's robustness
In this part we are trying to identify whether the model that we have already tested are robust or if a lack of robustness could explain some strange results that we obtained. 

```{r}
cat("=== VIF: Social Model ===\n")
vif_social <- vif(fit) 
print(vif_social)
cat("\n") 

cat("=== VIF: Violence Model ===\n")
vif_violence <- vif(fit_violence)
print(vif_violence)
cat("\n")

cat("=== VIF: Decision Model ===\n")
vif_decision <- vif(fit_decision)
print(vif_decision)
cat("\n")
```
In our three models, multicollinearity appears to be moderate, as all VIF values fall between 1 and 5. Nevertheless, this indicates that some degree of multicollinearity is present, suggesting that the models we tested are not entirely free from correlation among the predictors. The last model appears to be more robust, as it exhibits lower multicollinearity compared to the others.

**Breusch-Pagan test (heteroscedasticity)**

```{r}
cat("=== Breusch-Pagan test : Social Model ===\n")
bptest(fit)

cat("=== Breusch-Pagan test : Violence Model ===\n")
bptest(fit_violence)

cat("=== Breusch-Pagan test : Decision-making Model ===\n")
bptest(fit_decision)
```

To validate the reliability of our statistical inferences, we performed
the Studentized Breusch-Pagan test on our three models. For the Social
Independence (p=0.0016), Attitude toward Violence (p\<2.2e−16), and
Decision-Making (p\<1.9e−06) models, the null hypothesis of
homoscedasticity was strongly rejected.

This result confirms that the variance of the residuals is not constant
across observations, which is common in cross-country analyses where
heteroscedasticity often arises from differences in country size or
institutional volatility. To address this, we systematically applied
White’s Heteroscedasticity-Consistent (HC1) standard errors. This robust
estimation ensures that our t-statistics and p-values remain valid and
that the reported significance of our predictors—such as GDP and
financial inclusion—is not artificially inflated.

**Cook test**

```{r}
par(mfrow = c(1, 3))
par(mar = c(4, 4, 3, 1))
cooks_social <- cooks.distance(fit)
plot(1:length(cooks_social), cooks_social, type = "h", 
     xlab = "Observation index", ylab = "Cook's Distance",
     main = "Cook's Distance for Social Model")
abline(h = 1, col = "red", lty = 2)

cooks_violence <- cooks.distance(fit_violence)
plot(1:length(cooks_violence), cooks_violence, type = "h", 
     xlab = "Observation index", ylab = "Cook's Distance",
     main = "Cook's Distance for Violence Model")
abline(h = 1, col = "red", lty = 2)


cooks_decision <- cooks.distance(fit_decision)
plot(1:length(cooks_decision), cooks_decision, type = "h", 
     xlab = "Observation index", ylab = "Cook's Distance",
     main = "Cook's Distance for Decision Model")
abline(h = 1, col = "red", lty = 2)

```
Cook’s Distance analysis shows that none of the observations in our three models are highly influential, as all values remain well below the common threshold of 1.


**Model Reliability and Predictive Power Statistics**
```{r}
get_model_stats <- function(model) {
  s <- summary(model)
  data.frame(
    R2 = s$r.squared,
    Adj_R2 = s$adj.r.squared,
    F_statistic = s$fstatistic[1],
    p_value_F = pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE)
  )
}


model_performance <- bind_rows(
  data.frame(Model = "Social Independence", get_model_stats(fit)),
  data.frame(Model = "Attitude to Violence", get_model_stats(fit_violence)),
  data.frame(Model = "Decision-Making", get_model_stats(fit_decision))
)

knitr::kable(model_performance, 
             digits = 3, 
             row.names = FALSE, # Supprime les numéros de ligne inutiles à gauche
             caption = "Global Model Performance Indicators",
             col.names = c("Model Domain", "R²", "Adjusted R²", "F-Statistic", "p-value (F)"))
```

 

**Limits of our analysis** : 
A major limitation of our analysis is the selection bias inherent in the available datasets. Our graphical and econometric results primarily reflect countries where data collection is institutionalized. Low-income countries and fragile states are significantly underrepresented due to missing values in the Findex or WHO surveys. Therefore, our findings may not be fully generalizable to the most vulnerable populations, where the barriers to empowerment are likely even more structural and severe than those observed in our sample. Furthermore since our analysis relies in cross-sectional data, it limits causal interpretation and misses long term trends. Finally the ability of our different models to explain women's empowerment differ for each type of empowerment that we observe, for instance our model fail to explain efficiently women's empowerment in the social or attitude to violence domains.

A faire :

description graphique géographique, rédiger la conclusion





## Conclusion



## Bibliography 

"The roots of Gender Inequality in Developing Countries" Department of Economics, Northwestern University, Evanston, Illinois S.Jayachandran (2015)

"Household Matters: Revisiting the Returns to Capital among Female Microentrepreneurs" Bernhardt et al AMERICAN ECONOMIC REVIEW (2019) 

"The Quiet Revolution That Transformed Women's Employment, Education, and Family" Claudia Goldin AMERICAN ECONOMIC REVIEW (2006)





