---
title: "Determinants of Women's Empowerment: A Cross-Country Analysis"
author: "Betty Krebs & Marilys Guiban"
format: 
  pdf:
    echo: false   
    message: false
    warning: false
    toc: true
    number-sections: true
    colorlinks: true
    geometry:
      - margin=1in
editor: 
  markdown: 
    wrap: 72
---

*This report was generated using Quarto and RStudio. All statistical
analyses were performed in R, and the final document was compiled into
PDF format using the TinyTeX distribution to ensure high-quality
typesetting of econometric tables and LaTeX formulas.*

```{r}
# install.packages("WDI")  
# install.packages("sandwich") 
# install.packages("lmtest")
library(here)
library(ggplot2)
library(dplyr)
library(vroom)
library(tidyr)
library(knitr)
library(stringr)
library(WDI)
library(lmtest)
library(sandwich)
library(car)
```

## Research question presentation

The empowerment of women is tackled by the United Nations as a key step towards development as it holds the fifth position among the Sustainable Development Goals (SDGs), just behind the objectives of no poverty, zero hunger, good health and quality education. 

Yet, as of January 2025, women held only 27.2% of national parliament seats. This represents a tiny 0.3 point increase compared to 2024. 

The reviews of literature show that the progress is even slower in the financial domain, thus hindering women's empowerment. Scholars stress the need for more financial independence for women. For instance, Sujan (2016) makes a list of financial tasks essential in a household that could be endorsed by women, Bhumija Chauhan (2024) shows there is a robust evidence of a correlation between financial independence and economic, social and political empowerment. Pal, Gupta and Joshi(2022) show that earning status, social security schemes, and financial decision-making at household level are significant to promote women empowerment through financial inclusion. 

**Transition missing**

Research question : Do the quality of institutions and the development
of institutional financing systems enable women to empower themselves in
the social and decision-making domains ?

The aim of our research question is to understand whether countries with
higher-governance and more developed financial inclusion infrastructures
provide better opportunities for women to participate in economic,
social and decision-making activities. By combining multiple datasets,
our study integrates information on financial behavior, governance
indicators and women's empowerment to explore this relationship in a
comprehensive manner.

Dependent variables : women empowerment in the social domain and in the
decision making domain.

Explanatory variables : From the Findex dataset : borrowing from family
and friends (borrow_family_friends), using digital payments
(digital_payment), saving in a bank account (save_bank_account),
borrowing to start a business (borrow_start_business), and having a bank
account (has_bank_account)

From the Worldwide governance index : Voice and Accountability: Estimate
and Voice and Accountability: Percentile Rank, Rule of Law: Estimate and
Rule of Law: Percentile Rank, and Political Stability and Absence of
Violence/Terrorism: Estimate and Political Stability and Absence of
Violence/Terrorism: Percentile Rank. 

Our analysis relies primarily on three datasets. First the Global Findex
Database 2025, which privides demand-side data on financial inclusion,
including indicators such as account ownership, saving behaviour,
borrowing formal and informal sources and use of digital financial
services. These variables are critical to understanding women's access
to financial resources and their ability to manage risk and participate
in economic activity.

The WGI dataset offers a measure of institutional quality across
multiple dimensions such as political stability, rule of law, regulatory
quality and voice and accountability. These indicators allow us to
assess the institutional context in where financial systems operate.

The Empowerment dataset proposes an estimation of the empowerment of women in the social independence domain, the decision-making model and the attitude to violence.

To provide a more nuanced analysis, we will do distinctions between
countries to see how cultural and legal frameworks often influence
women's access to financial resources and participation in economic
activity.

In terms of methodology our project will combine descriptive statistics,
correlation analysis and regression models. The goal is to assess the
relationship between governance quality, financial inclusion and women's
empowerment. The clean datasets allow us to focus on the most relevant
variables.

In our analysis we will focus on the most recent data and give
historical analysis with previous litterature made on this question.

To answer this research question wo will : 
1° do a comparative geographical analysis of countries per GDP (low, middle, high income country) and dominant religion (Muslim vs non Muslim), and analyze women empowerment through scatter points clouds. 
2° do an analysis of the correlation between finance institutions and women empowerment
3° realise econometric regressions by isolating financial inclusion and level of autonomy for women. The tests will focus on the robustness of the coefficients and the quality of the model.

## Dataset description

### Empowerment

```{r}
Empowerment <- readRDS(here("datafinalproject", "Empowerment_clean.rds"))
```

This dataset comes from the World Health Organization, which issues the
survey-based women's empowerment index (SWPER). This index was first
tested in Africa in 2017 and was then extended to the rest of the world.
It is the first index in the gender domain to propose an
individual-level comparison between a large number of countries and over
time.

From this survey are drawn 9 indicators, distributed across 3 domains:
    - *Social Independence Domain* 
    - *Decision-Making Domain*
    - *Violence Domain*

### Findex (cleaned dataset)

```{r}
Findex <- readRDS(here("datafinalproject", "Findexclean.rds"))
```

In the cleaned format of this dataset we have 8 variables : 

-   Country, codecountry and year : giving informations about the observations

-   "has_bank_account" : The percentage of respondents who report being
    employed in the private sector and receiving any money from their
    employer in the past year in the form of a salary or wages for doing
    work, and who received that money directly into a bank or similar
    financial institution account

-   "borrow_start_business" : The percentage of respondents who report
    borrowing any money to start or operate a business in the past year.

-   "save_bank_account" : The percentage of respondents who report
    saving or setting aside any money at a bank or similar financial
    institution or using a mobile money account to save in the past
    year.

-   "digital payment" : The percentage of respondents who report using
    mobile money, a debit or credit card, or a mobile phone to make a
    payment from an account--or report using the internet to pay bills
    or to buy something online or in a store--in the past year. This
    includes respondents who report paying bills, sending or receiving
    remittances, receiving payments for agricultural products, receiving
    government transfers, receiving wages, or receiving a public sector
    pension directly from or into a bank or similar financial
    institution account or through a mobile money account in the past
    year.

-   "borrow family_friends" : The percentage of respondents who report
    borrowing any money from family, relatives, or friends in the past
    year

### WGI (cleaned format)

```{r}
WGI <- readRDS(here("datafinalproject", "WGI_cleaned.rds"))
```

For our research question we are using : 

-   political stability and the absence of terrorism as an explicative variable of women empowerment.      It measures perceptions of the likelihood of political instability and/or politically-motivated        violence, including terrorism. Estimate gives the country's score on the aggregate indicator, in       units of a standard normal distribution, i.e. ranging from approximately -2.5 to 2.5.

-   Voice and accountability : captures perceptions of the extent to
    which a country's citizens are able to participate in selecting
    their government, as well as freedom of expression, freedom of
    association, and a free media. Estimate gives the country's score on
    the aggregate indicator, in units of a standard normal distribution,
    i.e. ranging from approximately -2.5 to 2.5.

-   rule of law : Regulatory Quality captures perceptions of the ability
    of the government to formulate and implement sound policies and
    regulations that permit and promote private sector development.

For each indicators we keep the estimate and the percentile value: 
-Estimate gives the country's score on the aggregate indicator, in units
of a standard normal distribution. 
- Percentile rank indicates the country's rank among all countries covered by the aggregate indicator,
with 0 corresponding to lowest rank, and 100 to highest rank. Percentile
ranks have been adjusted to correct for changes over time in the
composition of the countries covered by the WGI.

### Joined file

The final dataset integrates all variables from the beforementioned
sources. By using the country code and year as primary keys, we
successfully merged the data into a single, comprehensive table for our
analysis.

```{r}
final_cleaned_all <- readRDS(here("datafinalproject", "final_cleaned_all.rds"))
```

To improve analytical precision, the final_recent dataset was restricted
to 2023 empowerment values. This minimizes the time gap with the 2024
WGI and Findex data, ensuring that the relationship between governance
and empowerment is based on nearly contemporaneous observations.

```{r}
final_recent <- readRDS(here("datafinalproject", "final_recent.rds"))
```

```{r}
final_with_gdp <- readRDS(here("datafinalproject", "final_with_gdp.rds"))
```


## Graphical representations

In this part we will do graphical analysis to understand what is
significant in women's empowerment. We decided not to analyse high
income countries since our research question explores development issues
and also due to a lack of data.

### Temporal analysis : how women's empowerment has evolved these past years ?

```{r}
final_cleaned_all |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  filter(!is.na(wbincome2024), wbincome2024 != "High income") |>
  group_by(year, wbincome2024, domaine) |> 
  summarise(mean_emp = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(mean_emp)) |> 
  ggplot(aes(x = year, y = mean_emp, color = wbincome2024, group = wbincome2024)) +
  geom_line(linewidth = 0.8, alpha = 0.6) + # 'linewidth' remplace 'size' pour les lignes
  geom_point(size = 1.5) + 
  facet_wrap(~domaine) +
  labs(
    title = "Historical Evolution of Women's Empowerment (2010-2023)",
    subtitle = "Trends by World Bank Income Group",
    x = "Year", 
    y = "Average Score (%)",
    color = "Income Group"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

The historical data (1999-2023) reveals an overall upward trend in
women's empowerment, particularly in the Attitude to Violence and
Decision-Making domains across all income groups. However, this progress
is characterized by significant volatility, suggesting that empowerment
is not a linear process but one sensitive to periodic socio-economic
shocks. Notably, the Social Independence domain shows a concerning
stagnation, or even a slight decline in some Upper-Middle and Low-Income
contexts, indicating that physical mobility and financial autonomy
remain the most difficult barriers to break, despite improvements in
other areas.

### Analysis of empowerment by income group

```{r}
final_recent |>
  filter(!is.na(estimate), !is.na(wbincome2024), wbincome2024 != "High income") |> 
  group_by(wbincome2024, country) |> 
  summarise(score_empowerment = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  ggplot(aes(x = wbincome2024, y = score_empowerment, fill = wbincome2024)) +
  geom_boxplot() + 
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal()
```

The distribution of empowerment scores reveals a significant shift
across income groups. While Low-Income countries struggle with high
inequality and low averages, Lower-Middle-Income nations show a
consolidated progress with more homogeneous and higher scores. In
contrast, Upper-Middle-Income countries exhibit a dual reality: despite
a higher ceiling of empowerment, the low positioning of the mean within
a tall box suggests persistent internal disparities, where a large
segment of the female population has yet to catch up with the top
achievers.

Let's analyse this deeper by focusing on social domain and decision
making domain by countries :

```{r}
final_recent |>
  filter(indicator_name %in% c(
    "Women with high empowerment in the attitude to violence domain (%)",
    "Women with high empowerment in the decision-making domain (%)",
    "Women with high empowerment in the social independence domain (%)"
  ), wbincome2024 != "High income") |>
  # We add a security filter pour empty scores
  filter(!is.na(estimate)) |> 
  group_by(country, wbincome2024, indicator_name) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  ggplot(aes(x = wbincome2024, y = score_moyen, fill = wbincome2024)) +
  # na.rm = TRUE suppresses the warning on the deleted lines
  geom_boxplot(outlier.shape = NA, alpha = 0.7, na.rm = TRUE) +
  facet_wrap(~domaine, scales = "free_y") +
  labs(
    title = "Empowerment Dimensions by Income Group",
    x = "Income Category",
    y = "Average Score (%)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  )
```

A cross-sectional analysis by income group reveals heterogenous patterns
across the three empowerment domains. In the Attitude toward Violence
domain, a clear positive correlation with national wealth is observed as
the median score rises from 52% in Low-Income countries to 75% in
Upper-Middle-Income contexts. However, the widening of the interquartile
range in wealthier groups suggests that economic growth alone does not
harmonize social norms.

In contrast, the Decision-Making domain shows extreme volatility within
the Low-Income group, with scores ranging from 23% to 63%. This
'long-box' phenomenon indicates that domestic authority can reach high
levels even in resource-constrained environments, likely driven by
specific institutional or cultural factors. Finally, Social Independence
appears to be the most rigid dimension, characterized by lower overall
scores and narrower distributions. The transition from a 23% median in
Low-Income countries to only 60% in Upper-Middle countries confirms that
physical and financial autonomy faces a 'structural ceiling' that is
significantly more resistant to economic development than other
dimensions of empowerment.

### Analysis of empowerment by the quality of institutions

In order to do this we will analyse graphically whether there seems to be
a correlation between empowerment and the quality of institutions.

*Political stability and women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, political_stability) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(political_stability), !is.na(score_moyen)) |>
  ggplot(aes(x = political_stability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Does Political Stability Foster Women's Empowerment?",
    subtitle = "Comparison across economic levels (Low to Upper-Middle Income)",
    x = "Political Stability Index (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```

The divergent trends across income groups suggest that the relationship
between political stability and empowerment is not linear but
context-dependent. While stability is a fundamental prerequisite for
empowerment in Low-Income settings, its impact diminishes or even
reverses in Lower-Middle-Income contexts. In contrast, for
Upper-Middle-Income nations, the positive correlation in decision-making
indicates that stability provides the necessary environment for highly
educated women to consolidate their role in governance and household
management.

*Voice and accountability and Women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, voice_accountability) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(voice_accountability), !is.na(score_moyen)) |>
  ggplot(aes(x = voice_accountability, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Correlation between Voice and Accountability and Women's Empowerment",
    x = "Voice & Accountability (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal()
```

Voice and Accountability shows a positive correlation with reduced
violence across most groups, yet reveals a striking divergence in
autonomy: while Upper-Middle-Income countries see gains across all
domains, Low and Lower-Middle-Income nations experience a 'participation
paradox' where increased political voice coincides with a decline in
domestic decision-making and social independence.

*Rule of law and women empowerment*

```{r}
final_recent |>
  group_by(country, wbincome2024, domaine, rule_of_law) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(rule_of_law), !is.na(score_moyen)) |>
  ggplot(aes(x = rule_of_law, y = score_moyen, color = wbincome2024)) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.2) + 
  facet_wrap(~domaine, ncol = 3) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Correlation between rule of law and Women's Empowerment",
    subtitle = "Analysis of legal frameworks across income groups",
    x = "Rule of Law (Percentile Rank)",
    y = "Average Empowerment Score (%)",
    color = "Income Group"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = NA),
    strip.text = element_text(face = "bold"),
    panel.spacing = unit(1, "lines")
  )
```

Rule of Law acts as a universal driver for social independence and
attitudes against violence across all income groups; however, it fails
to penetrate the domestic sphere in lower-income economies, where
decision-making power continues to decline despite strengthening legal
frameworks.

### Analysis of empowerment by development of financing systems

*Digital payments*

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, digital_payment) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(digital_payment), !is.na(score_moyen)) |>
  
  ggplot(aes(x = digital_payment, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the development of digital payment on Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population with Digital Payments",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The data on digital financial inclusion reveals a complex landscape:
while it serves as a catalyst for changing attitudes toward violence in
Lower-Middle-Income countries, it shows a global negative correlation
with decision-making and social independence, suggesting that digital
tools alone—potentially due to data scarcity or the 'digital
divide'—have yet to translate into broader structural autonomy for women
in developing economies.

*Is having a bank account correlated with women empowerment ?*

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, has_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(has_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = has_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population which has a bank accont and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who has a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

Analysis of formal bank account ownership reveals a transitional stage:
while it fosters social independence and decision-making power in
Lower-Middle-Income countries, it shows a slight negative correlation
with attitudes toward violence globally and a decline in decision-making
for Upper-Middle-Income groups, potentially reflecting a shift where
basic banking is no longer a sufficient lever for empowerment in more
advanced economies.

*Ability to borrow to bank to start a business*

```{r}
final_cleaned_all |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, borrow_start_business) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(borrow_start_business), !is.na(score_moyen)) |>
  
  ggplot(aes(x = borrow_start_business, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population which borrow to start its business and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who borrow to start a business",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The data analysis reveals that while borrowing to bank to start a
business correlate with increased social independence in
Lower-Middle-Income countries, they show a global decline in
decision-making power and attitudes toward violence, suggesting that
digital financial tools may increase women's mobility without yet
shifting deep-seated domestic power dynamics or social norms.

```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    ),
    wbincome2024 != "High income"
  ) |>
  mutate(domaine = str_remove_all(indicator_name, "Women with high empowerment in the | domain \\(%\\)")) |>
  group_by(country, wbincome2024, domaine, save_bank_account) |>
  summarise(score_moyen = mean(estimate, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(save_bank_account), !is.na(score_moyen)) |>
  
  ggplot(aes(x = save_bank_account, y = score_moyen, color = wbincome2024)) +
  geom_count(alpha = 0.3, size = 1) + 
  geom_smooth(method = "lm", formula = y ~ x, se = TRUE, aes(fill = wbincome2024), alpha = 0.1) +
  facet_wrap(~domaine) +
  scale_color_brewer(palette = "Set1") +
  scale_fill_brewer(palette = "Set1") +
  labs(
    title = "Correlation between the share of population who save on a bank account and Women's Empowerment",
    subtitle = "Size of dots represents the number of countries; Shaded areas show confidence intervals",
    x = "% of population who saves of a bank account",
    y = "Empowerment Score (%)",
    size = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", panel.grid.minor = element_blank())
```

The analysis of formal savings shows that while it correlates with
increased attitudes against violence in Lower-Middle-Income countries,
it appears decoupled from broader empowerment; in fact, formal savings
coincide with a decline in decision-making and social independence
across most groups, suggesting that saving may be a protective financial
strategy rather than a tool for immediate domestic or social autonomy.

Limits of our analysis : we can observe that there is a lack of data on
low income country it could be due to


##Empowerment in Muslim vs non-Muslim countries 

We determined a list of Muslim countries following the CIA World Factbook 
establishing a list of countries where a majority (more than 50%) of Muslims 
lives. This list does not count Mayotte as it considered a French territory. 
Palestine is considered by its two halves, the Gaza strip and West Bank. 

```{r}
muslim_countries <- c(
  "SAU", "IRN", "IRQ", "EGY", "MAR", "DZA", "TUN", "JOR", "PAK", "BGD",
  "IDN", "TUR", "AZE", "QAT", "KWT", "OMN", "YEM", "AFG", "MLI", "NER", "SOM", 
  "MRT", "MDV", "COM", "UZB", "DJI", "SEN", "LBY", "TJK", "GMB", "TKM", "SYR", 
  "SDN", "KGZ", "GIN", "ALB", "SLE", "BRN", "KAZ", "BHR", "MYS", "BFA", "LBN", 
  "TCD", "NGA", "BIH"
)
final_cleaned_all |>
  mutate(muslim_country = if_else(codecountry %in% muslim_countries,
                              "Muslim countries",
                              "Non-Muslim countries")
  )
```


```{r}
final_cleaned_all |>
  mutate(
    domaine = str_remove_all(
      indicator_name,
      "Women with high empowerment in the | domain \\(%\\)"
    )
  ) |>
  filter(domaine == "decision-making") |>
  group_by(codecountry, muslim_country) |>
  summarise(
    score_moyen = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = muslim_country, y = score_moyen, fill = muslim_country)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 1) +
  labs(
    title = "Women's empowerment in the decision-making domain",
    x = "",
    y = "Share of women with high empowerment (%)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

```{r}
final_recent |>
  group_by(muslim_countries, domaine) |>
  summarise(
    mean_empowerment = mean(empowerment, na.rm = TRUE),
    sd_empowerment = sd(empowerment, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  ) |>
  ggplot(aes(x = muslim_countries, y = mean_empowerment, fill = muslim_countries)) +
  geom_col(width = 0.6) +
  geom_errorbar(
    aes(ymin = mean_empowerment - sd_empowerment,
        ymax = mean_empowerment + sd_empowerment),
    width = 0.2
  ) +
  facet_wrap(~ domaine, scales = "free_y") +
  labs(
    title = "Average women's empowerment by domain and country group",
    x = "",
    y = "Mean empowerment index"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold")
  )

```

### Geographical analysis 


```{r}
final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    )
  ) |>
  mutate(
    domaine = str_remove_all(
      indicator_name,
      "Women with high empowerment in the | domain \\(%\\)"
    )
  ) |>
  group_by(whoreg6, domaine) |>
  summarise(
    mean_empowerment = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  ggplot(aes(x = domaine, y = whoreg6, fill = mean_empowerment)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "C",
    name = "Mean empowerment (%)"
  ) +
  labs(
    title = "Average women's empowerment by region and domain",
    x = "Empowerment domain",
    y = "World Health Organisation region"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 20, hjust = 1),
    plot.title = element_text(face = "bold")
  )

```

```{r}
library(tidyr)

df_dumbbell <- final_recent |>
  filter(
    indicator_name %in% c(
      "Women with high empowerment in the attitude to violence domain (%)",
      "Women with high empowerment in the decision-making domain (%)",
      "Women with high empowerment in the social independence domain (%)"
    )
  ) |>
  mutate(
    domaine = str_remove_all(
      indicator_name,
      "Women with high empowerment in the | domain \\(%\\)"
    )
  ) |>
  group_by(whoreg6, domaine) |>
  summarise(
    mean_empowerment = mean(estimate, na.rm = TRUE),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = domaine, values_from = mean_empowerment)

ggplot(df_dumbbell) +
  geom_segment(
    aes(
      y = whoreg6,
      yend = whoreg6,
      x = `attitude to violence`,
      xend = `decision-making`
    ),
    color = "grey70",
    linewidth = 1
  ) +
  geom_point(
    aes(x = `attitude to violence`, y = whoreg6),
    color = "#1b9e77",
    size = 3
  ) +
  geom_point(
    aes(x = `decision-making`, y = whoreg6),
    color = "#d95f02",
    size = 3
  ) +
  labs(
    title = "Gap between empowerment domains by WHO region",
    x = "Mean empowerment (%)",
    y = "WHO region"
  ) +
  theme_minimal()

```


## Econometric regression with scatter plots

In this part we are using a dataset including the GDP of each country and we created a region.

To compute our different tests, we will do 4 hypothesis : 
1° **the ability of women to empower themselves in the social sphere** depends on
political freedom and their ability to access formal financial services
To test this we will use : 
- as a dependent variable : women empowerment in the social domain 
- as an explicative variable : voice_accountability, has_bank_account, digital_payments and
log_gdp

2° **The ability of women to empower themselves in violent situations**
To test this we will use 
- as a dependent variable : women attitude toward violence 
- as an explicative variable : log_gdp, political_stability, rule_of_law, save_bank_account

3° **The ability of women to empower themselves in the decision-making domain** 
To test this we will use : 
- as a dependent variable : women empowerment in the decision-making domain 
- as an explicative variable : log_gdp, voice_accountability, rule_of_law, borrow_start_business


**Test of women empowerment in social domain**

```{r}
fit <- lm(estimate ~ log_gdp + voice_accountability + digital_payment + has_bank_account, 
          data = final_with_gdp |> filter(domaine == "social independence"))

results_robust <- coeftest(fit, vcov = vcovHC(fit, type = "HC1"))

print(results_robust)
```

The empirical results reveal a complex relationship between economic
development, technology, and social empowerment:

-   **The Dominance of GDP in Development**: The coefficient for Log GDP
    per capita is positive and highly significant (β=19.05,p<0.001).
    This confirms that national wealth remains the primary engine for
    social independence, likely by increasing access to education,
    infrastructure, and health services for women.

-   **The effect of freedom and democracy** : the access to freedom and especially free speech             surprisingly does not allow for women empowerment in the social domain. On the contrary, it has a      significant negative effect on women empowerment (β=−0,124,p<0.01)

-   **The Digital Paradox**: The Digital Payment variable shows a
    significant negative coefficient (β=−22,736,p<0.05). This
    counter-intuitive result suggests that in the absence of targeted
    social policies, the mere expansion of digital financial tools does
    not automatically translate into greater social autonomy. It may
    reflect a "participation paradox" where digital tools are prevalent
    in societies where physical mobility remains restricted, or where
    digital surveillance offsets the benefits of financial inclusion.

**Attitude toward violence**

```{r}
fit_violence <- lm(estimate ~ log_gdp + rule_of_law + political_stability + save_bank_account, 
                   data = final_with_gdp %>% filter(domaine == "attitude to violence"))

results_violence <- coeftest(fit_violence, vcov = vcovHC(fit_violence, type = "HC1"))
print(results_violence)
```

```{r}
fit_decision <- lm(estimate ~ log_gdp + voice_accountability + borrow_start_business, 
                   data = final_with_gdp %>% filter(domaine == "decision-making"))

results_decision <- coeftest(fit_decision, vcov = vcovHC(fit_decision, type = "HC1"))
print(results_decision)
```

This model reveals three different things : 
- **the power of economic growth**: the effect of Log GDP is even stronger here (β=32.28) than in
the Social domain. This suggest that at the macro level, national wealth
is a prerequisite for women to gain agency in household
decision-making. 
- **The effect of Voice and accountability** : The
coefficient is negative and highly significant (β=−1.004). It suggests
that in countries with higher political freedom and citizen
participation, women's reported power in domestic decisions actually
tends to be lower. 
- **The Banker lever** : Unlike the "Digital Payment"
paradox found earlier, Having a Bank Account is a very strong positive
predictor (β=19.07).

### Testing model's robustness

```{r}
vif_social <- vif(fit) 
print(vif_social)

vif_violence <- vif(fit_violence)
print(vif_violence)

vif_decision <- vif(fit_decision)
print(vif_decision)
```

**Breusch-Pagan test (heteroscedasticity)**

```{r}
bptest(fit)
bptest(fit_violence)
bptest(fit_decision)
```

To validate the reliability of our statistical inferences, we performed
the Studentized Breusch-Pagan test on our three models. For the Social
Independence (p=0.0016), Attitude toward Violence (p\<2.2e−16), and
Decision-Making (p\<1.9e−06) models, the null hypothesis of
homoscedasticity was strongly rejected.

This result confirms that the variance of the residuals is not constant
across observations, which is common in cross-country analyses where
heteroscedasticity often arises from differences in country size or
institutional volatility. To address this, we systematically applied
White’s Heteroscedasticity-Consistent (HC1) standard errors. This robust
estimation ensures that our t-statistics and p-values remain valid and
that the reported significance of our predictors—such as GDP and
financial inclusion—is not artificially inflated.

**Cook test**

```{r}
plot(fit, which = 4)
plot(fit_violence, which = 4)
plot(fit_decision, which = 4)
```

**Model Reliability and Predictive Power Statistics**
```{r}
get_model_stats <- function(model) {
  s <- summary(model)
  data.frame(
    R2 = s$r.squared,
    Adj_R2 = s$adj.r.squared,
    F_statistic = s$fstatistic[1],
    p_value_F = pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE)
  )
}

stats_social <- get_model_stats(fit)
stats_violence <- get_model_stats(fit_violence)
stats_decision <- get_model_stats(fit_decision)

model_performance <- bind_rows(
  mutate(stats_social, Model = "Social"),
  mutate(stats_violence, Model = "Violence"),
  mutate(stats_decision, Model = "Decision")
)

knitr::kable(model_performance, digits = 3, caption = "Global Model Performance Indicators")
```

With this regression we are trying to see whether the impact of financial inclusion depend on national economic development.It seems that despite our analysis the R squared is still low. 
To explain this significant difference of the value of the R squared between our decision making model and empowerment in the social domain we can perhaps use some studies which show that social mobility and empowerment is governed by informal institutions. 

```{r}
fit_social_inter <- lm(estimate ~ log_gdp * has_bank_account + voice_accountability, 
                       data = final_with_gdp %>% filter(domaine == "social independence"))

results_inter <- coeftest(fit_social_inter, vcov = vcovHC(fit_social_inter, type = "HC1"))
print(results_inter)

summary(fit_social_inter)$adj.r.squared
```




## Diagnosis tests

A faire : 
- améliorer la forme du devoir : formulation de la question de recherche doit être mieux amenée, contextualisée, éventuellement en faisant appel à de la littérature scientifique
- mieux expliciter les résultats rigoureusement : faire attention dans les interprétations graphiques à ne pas faire d'analyse causales mais simplement des corrélations entre variables 
A un moment dans l'analyse graphique on a très peu de données sur les pays pauvres : expliciter pourquoi et dire que donc on a un biais dans notre analyse 
- pour chacun des résultats étonnant que l'on a dans les régressions ou interprétations graphiques trouver de la littérature scientifique et donner une analyse temporelle dans le temps de l'empowerment des femmes (si elle existe qui permettrait d'expliquer ces résutlats : exemple le digital paradox, expliquer le voice and accountability paradox sur
l'émancipation des femmes dans la sphère sociale)
- sur la partie régression économétrique : mettre une table pour chacune des variables utilisées avec min/max/ nombre d'observations, améliorer l'interprétation des régressions / OLS, faire l'interprétation des tests VIF (évalue la multicolinéarité) / Breusch Pagan (homoscédasticité) / test de Cook, ajouter une analyse des limites de notre travail, éventuellement améliorer la clarté de la présentation de nos résultats avec une table qui résume tout 
- ajouter un test de Khi 2
Eventuellement on peut ajouter d'autres régressions ou modifier les hypothèses initiales qui sous tendent les régressions, l'idée était de faire un premier travail puis d'ajuster, éventuellement ajouter au début une analyse géographique en ajoutant une variable région (déjà présente dans le dataset : final_with_gdp)






## Conclusion
